Sampling Strategies in Decisions from Experience

### 2022-02-23 manuscript.Rmd
Update Section 2 and 3

Importantly, the information that is provided about the choice alternatives can come in many forms.
For example, in *decisions from description* (DfD), inputs take the form of explicit and complete descriptions of all relevant properties of the choice alternatives.
Yet, in *decisions from experience* [DfE, @hertwigDecisionsExperienceEffect2004], inputs take the form of *samples*.
In this latter case, the properties of the choice alternatives are not described and therefore not known with certainty, but they must be inferred from the consequences (hereafter: outcomes) that occurred when the alternatives were chosen in similar past decisions.
Although these distinct input forms may in principle carry the same information about a given set of choice alternatives, behavioral decision research has so far produced a large body of papers indicating that---if a quite general perspective is taken---there are robust differences between the choices that are made in DfD and DfE, leading to the notion of the *description-experience gap* [@hertwigDescriptionexperienceGapRisky2009].
In one reading of the gap, empirical choice patterns indicate that in DfD people choose *as if* a rare outcome is given more weight in the evaluation of a choice alternative than would be objectively warranted by the principle of mathematical expectation; in turn, in DfE, ceteris paribus, people choose *as if* the same rare outcome is given less weight than would be warranted [e.g., @barronSmallFeedbackbasedDecisions2003; @hertwigDecisionsExperienceEffect2004; @weberPredictingRiskSensitivity2004; @erevAnomaliesForecastsDescriptive2017; see @wulffMetaanalyticReviewTwo2018, for a comprehensive meta-analytic review].
Hereafter, these choice patterns are referred to as the as-if overweighting and underweighting of rare outcomes, respectively.

This paper uses the as-if underweighting of rare outcomes as a starting point for an inquiry into the sampling process upfront choices in DfE.
Essentially, it discusses and demonstrates in a simulation how different strategies of sequentially generating outcomes and integrating them into samples (hereafter: sampling strategies) can produce distinct choice patterns in DfE---among them, the as-if underweighting of rare outcomes [see @hillsInformationSearchDecisions2010, for the original formulation of this claim].

The rest of the paper is therefore structured as follows.
In the first brief section, a formal definition of DfE is provided.
In the subsequent section, cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993] is reviewed as a data model that can be used to describe choices in DfE in terms of deviations from the principle of mathematical expectation.
In the section that follows, sampling strategies are treated in quite some detail.
<!-- Outline of the section -->
Thereafter, a simulation is presented in which the theoretical claims that are made in the preceding section are implemented in a computational model to simulate choice data in DfE.
The predictions of the model are summarized and fitted in CPT.
The paper closes with a summary and a conclusion.
Note that this paper, including the simulations and data analyses, is reproducible---materials and instructions can be found on the GitHub repository.

@@

Now, more formally, let the axiomatic definition of a prospect be that of a discrete random variable's distribution.
Let therefore the tuple $(\Omega, \mathcal{F}, P)$ be a probability space. 
Following the @kolmogorovFoundationsTheoryProbability1950 axioms, $\Omega$ is the sample space containing a finite set of elements
$$
\omega_i \in \{\omega_1, ..., \omega_n\} = \Omega
\; .
\tag{2}
$$ 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$, i.e., the power set $\mathcal{P}(\Omega)$ with each element $A \in P(\Omega) = \mathcal{F}$ being a subset of $\Omega$.^[<!--Open footnote-->Note that this paper considers only the case of finite sample spaces. The exact axiomatic definition of $\mathcal{F}$, which can be found in Kolmogorov [-@kolmogorovFoundationsTheoryProbability1950, pp. 2-3 and 14-15], also covers infinite sample spaces and is that of a $\sigma$-algebra. For the finite case, however, $\mathcal{P}(\Omega)$ can be considered as such [@georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.1].<!--Close footnote-->]
$P$ is a probability measure
$$
P: \mathcal{F} \to [0,1]
\; ,
\tag{3}
$$
which assigns each subset $A \in \mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.<!--^[The probabilities $p(\omega)$ are determined by the probability mass function $p : \Omega \to [0,1]$, which is used to construct the probability measures $P(A) = \sum_{\omega \in A} p(\omega)$ for all $A \in \mathcal{F}$.]-->
The random variable $X$ is the measurable function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{4}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ being a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the probability measure
$$
P_X : \mathcal{F'} \to [0,1]
\tag{5}
$$
with 
$$
P_X(A' \in \mathcal{F'}) := P(X^{-1} \ A') = P(\{\omega: X(\omega) = x \in A'\})
\tag{6}
$$
[see @kolmogorovFoundationsTheoryProbability1950, p. 21-22].^[<!--Open footnote-->See also @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1, for a comprehensive yet accessible introduction to random variables.<!--Close footnote-->]
That is, the set consisting of a single outcome $x$ can be treated as a subset $\{x\} \in \mathcal{F'}$ and we therefore obtain $p_X(x) = P(X(\omega) = x)$.
Note that the probability measure in (5) and (6), respectively, is a discrete random variable's distribution and therefore a prospect.
To summarize and simplify, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
C = \{P_{X_1}, ... , P_{X_k}\}
\tag{7}
$$
with each element $c \in C$ being a distribution $P_X$ of a random variable defined on a different probability space.
Accordingly, each $c$ can itself be considered a set of outcome-probability pairs $(x_i, p_X(x_i)) \in \{(x_1, p_X(x_1)), ..., (x_n, p_X(x_n))\}$.

However, rarely in peoples' daily life, inputs take the form of explicit descriptions of all outcomes-probability pairs.
In such cases, people would make decisions from description (DfD), since the information from which the mind learns about the prospects are complete descriptions thereof.
Yet, rather often, people make decisions from experience (DfE), where the mind can learn about the prospects only from a record of past experiences with them.
That is, people may have experienced the possible outcomes of a prospect when they themselves or others have made the same decision repeatedly in the past, and the record of experienced outcomes for this particular prospect is a rough description of a sample.
The mind may then infer from the relative frequency with which an outcome occurred in such a sample of past decisions its probability of occurrence following a future decision.

Now, more formally, let the sample of a prospect be a set of realizations of the respective random variable $X$ according to its distribution $P_X$.^[<!--Open footnote-->Note that this paper considers the counterfactual case where samples are random, although it has been emphasized that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. To account for non-random sampling, e.g., in the form of sampling biases [e.g., @fiedlerBewareSamplesCognitiveecological2000] or dynamic information structures [e.g., @cohenEffectPerceivedPatterns2021; @plonskyRelianceSmallSamples2015], the i.i.d-assumption would need to be dropped, which is out of the scope of this paper.<!--Close footnote-->]
So, for the choice between $1, ..., k$ prospects, the inputs to DfE are generated by a set $S$ of stochastic processes
$$
S = \{X_{1t}, ... , X_{kt}\}
\tag{8}
$$
with $t \in \{1, ..., N\}$ denoting a sequence of independent and identically distributed (i.i.d.) random variables and $N$ denoting the length of the respective sequence.
The sets of realizations that result from these stochastic processes may then be used to make an uncertain inference about the respective sets of outcome-probability pairs in (7).

Of course, the laws of large numbers [LLN, see, e.g., @kolmogorovFoundationsTheoryProbability1950] apply to DfE and it follows that the accuracy of such inferences from a sample to a prospect depends, ceteris paribus, on the size of the samples.
In fact, one explanation for the as-if underweighting of rare outcomes in DfE and the description-experience gap outlined in the introduction is that people rely on small samples, which, due to the skewness of the binomial distribution, cause the relative frequencies with which rare outcomes occur in a sample to be smaller than their objective probabilities [@hertwigDecisionsExperienceEffect2004; @rakowBiasedSamplesNot2008; see also @hertwigDecisionsExperienceWhy2010].
The invoked argument is then that what may appear as if people were underweighting rare outcomes relative to what would be warranted on the basis of their objective probabilities---following the principle of mathematical expectation---may in fact be consistent with an as-if overweighting relative to what would be warranted on the basis of the experienced probabilities [@foxDecisionsExperienceSampling2006].
Yet, it seems that the small size of samples in DfE provides just one and only a partial explanation for the description-experience gap [@hauDescriptionexperienceGapRisky2008; @hauDecisionsExperienceStatistical2010; @ungemachAreProbabilitiesOverweighted2009; see also @wulffMetaanalyticReviewTwo2018].
More so, despite the LLN, the gap need not narrow at all, even for ever increasing sample sizes that are fully taken into account.
Particularly, this paper discusses and demonstrates in a simulation how different strategies of sequentially generating outcomes and integrating them into samples interact with the LLN to produce distinct choice patterns in DfE---including the as-if underweighting of rare outcomes---that tend to stabilize as the sample size increases.
Before so doing, however, the following section reviews how cumulative prospect theory (CPT) can be used as a data model that captures systematic deviations from the principle of mathematical expectation in the parameter estimates of a *value* and a *weighting function*.

@@

The choices people actually make between prospects---be that in DfD or DfE---are often described in terms of deviations from the principle of mathematical expectation, according to which the prospect with the largest expected value (EV)
$$
EV = \sum_i^n x_i \times p_X(x_i)
\tag{9}
$$ 
should be chosen (EV maximization).
To *describe* how people's actual choices deviate from this principle [see @erevAnomaliesForecastsDescriptive2017, for a recent replication of classical demonstrations],^[<!--Open footnote-->@tverskyAdvancesProspectTheory1992 addressed the *fourfold pattern of risk attitudes*---i.e., choices between prospects indicate risk aversion for gains of high probability and losses of low probability, and risk seeking for gains of low probability and losses of high probability---and related effects, e.g., the *certainty effect* and the *reflection effect*, that were already addressed by CPT's predecessor, prospect theory [@kahnemanProspectTheoryAnalysis1979]. Together, these empirical choice patterns indicate a violation of the principle of mathematical expectation and the expected utility (EU) principle [see @bernoulliExpositionNewTheory1954\/1738].<!--Close footnote-->]
the models of CPT [@tverskyAdvancesProspectTheory1992] and similar rank-dependent models [see @stottCumulativeProspectTheory2006, for an overview of models] fit choice data by maximizing^[<!--Open footnote-->Note that CPT may be amended by a stochastic choice rule to account for the probabilistic nature of choices [@rieskampProbabilisticNaturePreferential2008]. Most choice rules [@stottCumulativeProspectTheory2006, for an overview] are functions of the prospects' values $V$ and make the choice of a prospect more probable the greater its $V$ relative to those of other prospects.<!--Close footnote-->]
the value
$$
V = \sum_i^n v(x_i) \times \pi_i
\; ,
\tag{10}
$$ 
with objective outcomes $x_i$ being transformed by a value function $v$, and *cumulative decision weights* $\pi_i$ being determined by the difference between transformed cumulative probabilities of the distribution $P_X$.
More specifically, following @tverskyAdvancesProspectTheory1992, the objective outcomes are transformed by a value function
$$
\begin{aligned}
  v : \Omega' &\to \mathbb{R} \\
  x &\mapsto 
  \begin{cases}
     x_i^\alpha &\forall x_i \geq 0, \\
     -\lambda \times |x_i|^\alpha &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{11}
$$
with $\alpha \in [0,1]$ determining the degree of the function's concavity (convexity) over the positive (negative) outcome interval, and $\lambda > 1$ increasing the function's slope over the negative outcome interval only.
Each subjective value $v(x)$ is then multiplied (or: weighted) with a cumulative decision weight that takes the form
$$
\pi_i =
  \begin{cases}
     w^+(P(X \geq x_i)) - w^+(P(X > x_i)) \quad \forall x_i \geq 0, \\
     w^-(P(X \leq x_i)) - w^-(P(X < x_i)) \quad \text{else}
     \; , 
  \end{cases}
\tag{12}
$$
where $w$ is a monotonic increasing, nonlinear weighting function $w: [0,1] \to [0,1]$ satisfying $w^+(0) = w^-(0) = 0$ and $w^+(1) = w^-(1) = 1$.

Essentially, the proposed model is such that the cumulative decision weights derived for the outcomes may be greater or smaller than their objective probabilities, causing the transformed outcomes---i.e., the subjective values---to be over- or underweighted *in* CPT, respectively.
<!--Accordingly, one may adopt an as-if weighting terminology, stating that people choose *as if* they maximized the value in (10) and applied the weighting pattern that was estimated in CPT.-->
Now, several consequences for the weighting pattern that can be estimated in CPT follow from (12), which are reviewed one by one next:

@@

* The value of a cumulative decision weight depends on the estimated shape of the graph of $w$, the weighting function's parameters respectively.
That is, the shape of the graph of $w$ displays in which interval on the cumulative probability scale $[0,1]$ the images of $w$ take values that are greater or smaller than the respective cumulative probabilities---i.e., the preimages---and how much greater or smaller these image values are.
Hence, assuming a nonlinear graphical shape of $w$, the same numerical difference between objective cumulative probabilities may translate to cumulative decision weights of different value.

@@

To begin with, the gray-dashed identity lines in Figure\ \@ref(fig:w-shapes) imply a linear weighting pattern.
That is, since the images of $w$, $w(p)$, are equal to the respective cumulative probabilities $p$, all cumulative decision weights $\pi$ that are derived according to (12) would be equal to the differences between the respective objective cumulative probabilities and therefore also to the objective probabilities of the outcomes for which the weights are determined.
Such a linear weighting pattern is implied by $\gamma = 1$ and $\delta = 1$. 
Now, both deviations from $\gamma = 1$ and $\delta = 1$ produce a nonlinear graphical shape.
That is, for $\gamma > 1$ (green graphs), the graphs take a S-shape---running below (above) the identity line for small (large) cumulative probabilities---which is more accentuated for larger deviations from $\gamma = 1$.
Importantly, the same *small* numerical difference between two cumulative probabilities $p$ then translates to varying differences between the respective images $w(p)$ depending on the interval on the cumulative probability scale. 
Specifically, because of the S-shaped curvature, the differences between the images $w(p)$ may be smaller than those of the respective cumulative probabilities in the lower and upper part of the cumulative probability scale, but may be considerably greater in the middle part.
As a consequence, the cumulative decisions weights that are retrieved from the images of $w$ are smaller than the respective objective probabilities for small-probability outcomes of either a low rank or a high rank, but are larger for small-probability outcomes of middle ranks.^[<!--Open footnote-->Note that the distinction between outcomes of low/high rank and outcomes of middle rank is unnecessary for two-outcome prospects [see @tverskyAdvancesProspectTheory1992]. However, for prospects with finite many outcomes, this distinction is rather critical. E.g., the fact that CPT may then weigh outcomes of the same small probability differently, is one reason why the description-experience gap should not be equated with a reversed weighting pattern in CPT [see @hertwigConstructBehaviorGap2018]. I.e., neither could the gap then be interpreted in terms of an as-if underweighting (overweighting) of rare events in DfE (DfD), nor can any of the factors so far considered as a contributor to the gap (e.g., sample size, optional stopping, recency, etc.) plausibly explain why not all small-probability outcomes are affected in the same way.<!--Close footnote-->]
Note that for all $\gamma < 1$ (blue graphs) the graph of $w$ takes an inverse S-shape and the entire weighting pattern is reversed.
Moreover, the shape of the weighting function, and thus the weighting pattern, also depends strongly on $\delta$, which affects the overall elevation of the graph.
That is, for $\delta < 1$, the interval on the cumulative probability scale over which the weighting function runs below the identity line is greater than the interval over which the weighting function runs above the identity line; and this pattern reverses for $\delta > 1$.
More specifically, $\delta$ shifts across the entire cumulative probability scale the intervals within which small differences between objective cumulative probabilities translate to either smaller or greater cumulative decision weights.
In other words then, for any decrease (increase) in $\delta$, the underweighting of small probability outcomes extends to more outcomes at the lower (higher) end of the outcome range.

To summarize the preceding paragraphs, CPT can be used as a data model to describe choices between prospects by reducing systematic deviations from the principle of mathematical expectation to the parameter estimates of its value and weighting function.
Moreover, each estimated weighting function captures a weighting pattern.
Although the possible weighting patterns that can be captured depend on the exact parameterization of the weighting function [e.g., @goldsteinExpressionTheoryPreference1987], the latter's graph can usually take a more or less accentuated (inverse) S-shape.
The parameter estimates and the implied graphical shape of the weighting function then translate to a weighting pattern where small-probability outcomes of low or high rank are underweighted (overweighted) in the overall evaluation of a prospect according to (10).
Notably, the distinction between outcomes of low/high rank and outcomes of middle rank can be omitted for two-outcome prospects.
In such a case, the weighting pattern reduces to either an underweighting or an overweighting of rare outcomes in CPT. 
For the description of choices between at most two-outcome gambles, one may then adopt the as-if weighting terminology from above, by stating that people decide as if they were either underweighting or overweighting rare outcomes. 
Nevertheless, it cannot be concluded from the estimated weighting pattern that the mind does indeed perform any of the computations associated with the weighting of subjective values in CPT; rather, the mind processes the inputs in a way that the outputs, i.e., the actual choice patterns, approximate the solutions from the estimated CPT model [see, e.g., @griffithsRationalUseCognitive2015].
Therefore, this paper uses CPT merely as a data model to describe choices between prospects and to demonstrate how different sampling strategies can produce weighting patterns in CPT resembling those that were estimated for empirical choice data from DfE.

# Sampling Strategies

>>>>

Importantly, the information that is provided about the choice alternatives can come in many forms.
For example, in *decisions from description* (DfD), inputs take the form of explicit and complete descriptions of all relevant properties of the choice alternatives.
Yet, in *decisions from experience* [DfE, @hertwigDecisionsExperienceEffect2004], the properties of the choice alternatives are latent---i.e., not described---and therefore not known with certainty, but they must be inferred from the consequences (hereafter: outcomes) that occurred when the alternatives were chosen in similar past decisions.
Although these distinct input forms may in principle carry the same information about a given set of choice alternatives, behavioral decision research has so far produced a large body of papers indicating that---if a quite general perspective is taken---there are robust differences between the choices that are made in DfD and DfE, leading to the notion of the *description-experience gap* [@hertwigDescriptionexperienceGapRisky2009].
In one reading of the gap, empirical choice patterns indicate that in DfD people choose as if a rare outcome is given more weight in the evaluation of a choice alternative than would be objectively warranted by the principle of mathematical expectation; in turn, in DfE, ceteris paribus, people choose as if the same rare outcome is given less weight than would be warranted [e.g., @barronSmallFeedbackbasedDecisions2003; @hertwigDecisionsExperienceEffect2004; @weberPredictingRiskSensitivity2004; @erevAnomaliesForecastsDescriptive2017; see @wulffMetaanalyticReviewTwo2018, for a comprehensive meta-analytic review].
Hereafter, these choice patterns are referred to as the as-if overweighting and underweighting of rare outcomes, respectively.

This paper uses the as-if underweighting of rare outcomes as a starting point for an inquiry into the sampling process upfront choices in DfE.
Essentially, it discusses and demonstrates in a simulation how different strategies of sequentially generating outcomes and integrating them into *samples* (hereafter: sampling strategies) can produce distinct choice patterns in DfE---among them, the as-if underweighting of rare outcomes [see @hillsInformationSearchDecisions2010, for the original formulation of this claim].

The rest of the paper is therefore structured as follows.
In the first brief section, a formal definition of DfE is provided.
In the subsequent section, cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993] is reviewed as a data model that can be used to describe choices in DfE in terms of deviations from the principle of mathematical expectation.
In the section that follows, sampling strategies are treated.
<!-- Outline of the section -->
Thereafter, a simulation is presented in which the theoretical claims that are made in the preceding section are implemented in a computational model to simulate choice data in DfE.
The predictions of the model are summarized and fitted in CPT.
The paper closes with a summary and a conclusion.
Note that this paper, including the simulations and data analyses, is reproducible---materials and instructions can be found on the GitHub repository.

@@

Now, more formally, let the axiomatic definition of a prospect be that of a discrete random variable's distribution.
Let therefore the tuple $(\Omega, \mathcal{F}, P)$ be a probability space. 
Following the @kolmogorovFoundationsTheoryProbability1950 axioms, $\Omega$ is the sample space containing a finite set of elements, one of which results from choosing the respective prospect:
$$
\omega_i \in \{\omega_1, ..., \omega_n\} = \Omega
\; .
\tag{2}
$$ 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$, i.e., the power set $\mathcal{P}(\Omega)$ with each element $A \in P(\Omega) = \mathcal{F}$ being a subset of $\Omega$.^[<!--Open footnote-->Note that this paper considers only the case of finite sample spaces. The exact axiomatic definition of $\mathcal{F}$, which can be found in Kolmogorov [-@kolmogorovFoundationsTheoryProbability1950, pp. 2-3 and 14-15], also covers infinite sample spaces and is that of a $\sigma$-algebra. For the finite case, however, $\mathcal{P}(\Omega)$ can be considered as such [@georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.1].<!--Close footnote-->]
$P$ is a probability measure
$$
P: \mathcal{F} \to [0,1]
\; ,
\tag{3}
$$
which assigns each subset $A \in \mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.<!--^[The probabilities $p(\omega)$ are determined by the probability mass function $p : \Omega \to [0,1]$, which is used to construct the probability measures $P(A) = \sum_{\omega \in A} p(\omega)$ for all $A \in \mathcal{F}$.]-->
The random variable $X$ is the measurable function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{4}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ being a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the probability measure
$$
P_X : \mathcal{F'} \to [0,1]
\tag{5}
$$
with 
$$
P_X(A' \in \mathcal{F'}) := P(X^{-1} \ A') = P(\{\omega: X(\omega) = x \in A'\})
\tag{6}
$$
[see @kolmogorovFoundationsTheoryProbability1950, p. 21-22].^[<!--Open footnote-->See also @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1, for a comprehensive yet accessible introduction to random variables.<!--Close footnote-->]
That is, the set consisting of a single outcome $x$ is a subset $\{x\} \in \mathcal{F'}$ and we therefore obtain $p_X(x) = P(X(\omega) = x)$.
Note that the probability measure in (5) and (6), respectively, is a discrete random variable's distribution and therefore a prospect.
To summarize and simplify, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
C = \{P_{X_1}, ... , P_{X_k}\}
\tag{7}
$$
with each element $c \in C$ being a distribution $P_X$ of a random variable defined on its own probability space.
Accordingly, each $c$ can itself be considered a set of outcome-probability pairs $(x_i, p_X(x_i)) \in \{(x_1, p_X(x_1)), ..., (x_n, p_X(x_n))\}$.

However, rarely in peoples' daily life, inputs take the form of explicit descriptions of all outcomes-probability pairs.
In such cases, people would make DfD, since the information from which the mind learns about the prospects are complete descriptions thereof.
Yet, rather often, people make DfE, where the mind can learn about the latent prospects only from a record of past experiences with them.
That is, people may have experienced the possible outcomes of a prospect when they themselves or others have made the same decision repeatedly in the past, and the record of experienced outcomes for this particular prospect is a rough description of a sample.
The mind may then infer from the relative frequency with which an outcome occurred in such a sample of past decisions its probability of occurrence following a future decision.

Now, more formally, let the sample of a prospect be a set of realizations of the respective random variable according to its distribution $P_X$.^[<!--Open footnote-->Note that this paper considers random samples, although it has been emphasized that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. To account for non-random sampling, e.g., in the form of sampling biases [see @fiedlerBewareSamplesCognitiveecological2000] or dynamic information structures [see, e.g., @cohenEffectPerceivedPatterns2021; @plonskyRelianceSmallSamples2015], the i.i.d-assumption would need to be dropped.<!--Close footnote-->]
So, for the choice between $1, ..., k$ prospects, the inputs to DfE are generated by a set $S$ of stochastic processes
$$
S = \{X_{1t}, ... , X_{kt}\}
\tag{8}
$$
with $t \in \{1, ..., N\}$ denoting a sequence of independent and identically distributed (i.i.d.) random variables and $N$ denoting the length of the respective sequence.
The realizations that result from each stochastic process $s \in S$ may then be integrated into a sample of $N$ experienced outcomes or into multiple smaller samples.
The samples may then be used to make an uncertain inference about the latent prospects in (7), the sets of outcome-probability pairs respectively.

Of course, the laws of large numbers [LLN, see, e.g., @kolmogorovFoundationsTheoryProbability1950] apply to DfE and it follows that the accuracy of an inference from a sample about the properties of a latent prospect depends, ceteris paribus, on the size of the sample.
In fact, one explanation for the as-if underweighting of rare outcomes in DfE and the description-experience gap outlined in the introduction is that people rely on too small samples, which, due to the skewness of the binomial distribution, cause the relative frequencies with which rare outcomes occur in a sample to be smaller than their objective probabilities [@hertwigDecisionsExperienceEffect2004; @rakowBiasedSamplesNot2008; see also @hertwigDecisionsExperienceWhy2010].
The invoked argument is then that what may appear as if people were underweighting rare outcomes relative to what would be warranted on the basis of their objective probabilities---following the principle of mathematical expectation---may in fact be consistent with an as-if overweighting relative to what would be warranted on the basis of the experienced relative frequencies [@foxDecisionsExperienceSampling2006].
Yet, it seems that the small size of samples in DfE provides just one and only a partial explanation for the description-experience gap [@hauDescriptionexperienceGapRisky2008; @hauDecisionsExperienceStatistical2010; @ungemachAreProbabilitiesOverweighted2009; see also @wulffMetaanalyticReviewTwo2018].
More so, despite the LLN, the gap need not narrow at all, even for ever increasing sample sizes that are fully taken into account.
Particularly, this paper discusses and demonstrates in a simulation how different strategies of sequentially generating outcomes and integrating them into samples interact with the LLN to produce distinct choice patterns in DfE---including the as-if underweighting of rare outcomes---that tend to stabilize as the sample size increases.
Before so doing, however, the following section reviews how CPT can be used as a data model that captures systematic deviations from the principle of mathematical expectation in the parameter estimates of a *value* and a *weighting function*.

@@

The choices people actually make between prospects---be that in DfD or DfE---are often described in terms of deviations from the principle of mathematical expectation, according to which the prospect with the largest expected value (EV)
$$
EV = \sum_i^n x_i \times p_X(x_i)
\tag{9}
$$ 
should be chosen (EV maximization).
To *describe* how people's actual choices deviate from this principle [see @erevAnomaliesForecastsDescriptive2017, for a recent replication of classical demonstrations],^[<!--Open footnote-->@tverskyAdvancesProspectTheory1992 addressed the *fourfold pattern of risk attitudes*---i.e., choices between prospects indicate risk aversion for gains of high probability and losses of low probability, and risk seeking for gains of low probability and losses of high probability---and related effects, e.g., the *certainty effect* and the *reflection effect*, which were already addressed by CPT's predecessor, prospect theory [@kahnemanProspectTheoryAnalysis1979]. Together, these empirical choice patterns indicate a violation of the principle of mathematical expectation and the expected utility (EU) principle [see @bernoulliExpositionNewTheory1954\/1738].<!--Close footnote-->]
CPT and similar rank-dependent models [see @stottCumulativeProspectTheory2006, for an overview of models] fit choice data by maximizing^[<!--Open footnote-->Note that CPT may be amended by a stochastic choice rule to account for the probabilistic nature of choices [@rieskampProbabilisticNaturePreferential2008]. Stochastic choice rules [@stottCumulativeProspectTheory2006, for an overview] make the choice of a prospect more probable the better its valuation relative to those of other prospects.<!--Close footnote-->]
the value
$$
V = \sum_i^n v(x_i) \times \pi_i
\; ,
\tag{10}
$$ 
with objective outcomes $x_i$ being transformed by a value function $v$, and *cumulative decision weights* $\pi_i$ being determined by the difference between transformed cumulative probabilities of the distribution $P_X$.
More specifically, following @tverskyAdvancesProspectTheory1992, the objective outcomes are transformed by a value function
$$
\begin{aligned}
  v : \Omega' &\to \mathbb{R} \\
  x &\mapsto 
  \begin{cases}
     x_i^\alpha &\forall x_i \geq 0, \\
     -\lambda \times |x_i|^\alpha &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{11}
$$
with $\alpha \in [0,1]$ determining the degree of the function's concavity (convexity) over the positive (negative) outcome interval, and $\lambda > 1$ increasing the function's slope over the negative outcome interval only.
Each subjective value $v(x)$ is then multiplied (or: weighted) with a cumulative decision weight that takes the form
$$
\pi_i =
  \begin{cases}
     w^+(P(X \geq x_i)) - w^+(P(X > x_i)) \quad \forall x_i \geq 0, \\
     w^-(P(X \leq x_i)) - w^-(P(X < x_i)) \quad \text{else}
     \; , 
  \end{cases}
\tag{12}
$$
where $w$ is a monotonic increasing, nonlinear weighting function $w: [0,1] \to [0,1]$ satisfying $w^+(0) = w^-(0) = 0$ and $w^+(1) = w^-(1) = 1$.

Essentially, the proposed model is such that the cumulative decision weights derived for the outcomes may be greater or smaller than their objective probabilities, causing the transformed outcomes---i.e., the subjective values---to be over- or underweighted *in* CPT, respectively.
<!--Accordingly, one may adopt an as-if weighting terminology, stating that people choose *as if* they maximized the value in (10) and applied the weighting pattern that was estimated in CPT.-->
Now, several consequences for the weighting pattern that can be estimated in CPT follow from (12), which are reviewed next:

@@

* The value of a cumulative decision weight depends on the estimated shape of the graph of $w$, the weighting function's parameters respectively.
That is, the shape of the graph of $w$ displays in which interval on the cumulative probability scale $[0,1]$ the images of $w$, $w(p)$ take values that are greater or smaller than the respective cumulative probabilities and how much greater or smaller these images are.
Hence, assuming a nonlinear graphical shape of $w$, the same numerical difference between objective cumulative probabilities may translate to cumulative decision weights of different value.

@@

To begin with, the gray-dashed identity lines in Figure\ \@ref(fig:w-shapes) imply a linear weighting pattern.
That is, since the images of $w$, $w(p)$, are equal to the respective cumulative probabilities $p$, all cumulative decision weights $\pi$ that are derived according to (12) would be equal to the differences between the respective objective cumulative probabilities and therefore also to the objective probabilities of the outcomes for which the weights are determined.
Such a linear weighting pattern is implied by $\gamma = 1$ and $\delta = 1$. 
Now, both deviations from $\gamma = 1$ and $\delta = 1$ produce a nonlinear graphical shape.
That is, for $\gamma > 1$ (green graphs), the graphs take a S-shape---running below (above) the identity line for small (large) cumulative probabilities---which is more accentuated for larger deviations from $\gamma = 1$.
Importantly, the same *small* numerical difference between two cumulative probabilities $p$ then translates to varying differences between the respective images $w(p)$ depending on the interval on the cumulative probability scale. 
Specifically, because of the S-shaped curvature, the differences between the images $w(p)$ may be smaller than those between the respective cumulative probabilities in the lower and upper part of the cumulative probability scale, but may be considerably greater in the middle part.
As a consequence, the cumulative decisions weights are smaller than the respective objective probabilities for small-probability outcomes of either a low rank or a high rank, but they are larger for small-probability outcomes of a middle rank.^[<!--Open footnote-->Note that the distinction between outcomes of either low/high rank or middle rank is unnecessary for two-outcome prospects [see @tverskyAdvancesProspectTheory1992]. However, for finite many outcomes, the distinction is critical. More so, the fact that CPT may weigh outcomes of the same small probability differently, adds another reasons why the description-experience gap should not be equated with a reversed weighting pattern in CPT [see @hertwigConstructBehaviorGap2018]. I.e., neither could the gap then be interpreted in terms of an as-if underweighting (overweighting) of rare outcomes in DfE (DfD), nor would any of the potential contributors to the gap considered so far [see, e.g., @hertwigDescriptionexperienceGapRisky2009; @wulffMetaanalyticReviewTwo2018] treat some small-probability outcomes differently than others.<!--Close footnote-->]
Note that for all $\gamma < 1$ (blue graphs) the graph of $w$ takes an inverse S-shape and the entire weighting pattern is reversed.
Moreover, the shape of the weighting function, and thus the weighting pattern, depends strongly on $\delta$, which affects the overall elevation of the graph.
That is, for $\delta < 1$, the interval on the cumulative probability scale over which the weighting function runs below the identity line is greater than the interval over which the weighting function runs above the identity line; and this pattern reverses for $\delta > 1$.
More specifically, $\delta$ shifts across the entire cumulative probability scale the intervals within which small differences between objective cumulative probabilities translate to either smaller or greater cumulative decision weights.
In other words then, for any decrease (increase) in $\delta$, the underweighting of small probability outcomes extends to more outcomes at the lower (higher) end of the outcome range.

To summarize the preceding paragraphs, CPT is data model that can be used to describe choices between prospects by capturing systematic deviations from the principle of mathematical expectation in the parameter estimates of its value and weighting function.
Importantly, each estimated weighting function implies a weighting pattern.
Although the possible weighting patterns depend on the exact parameterization of the weighting function [e.g., @goldsteinExpressionTheoryPreference1987], the latter's graph usually takes a more or less accentuated (inverse) S-shape, which implies a weighting pattern where small-probability outcomes of low and high rank are underweighted (overweighted) in CPT's overall evaluation of a prospect according to (10).
Notably, for the case of choices between at most two-outcome prospects, the weighting pattern reduces to either an underweighting or overweighting of rare outcomes.
For the description of such choices, one may then adopt the as-if weighting terminology from above, by stating that people decide as if they were either underweighting or overweighting rare outcomes.
<!-- Add paragraph about CPT analysis of DfE-->
